---
title: "Wrangling with Tidyverse"
author: "J.Blondeau"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All the following question use the sttst_all_sites.xlsx file in the /data directory.


1. Import sttstj_all_sites and inspect the data set.  What is the data structure and the data type of each field?


```{r}
library(tidyverse)
library(readxl)
d <- read_xlsx("data/sttstj_all_sites.xlsx", sheet="locations")
class(d)
glimpse(d)
```

tibble
Columns: 5
$ yr_site <chr> "2001_241", "2001_243", "2001_244", "2001_247", …
$ year    <dbl> 2001, 2001, 2001, 2001, 2001, 2001, 2001, 2001, …
$ lat     <dbl> 18.31079, 18.34443, 18.34059, 18.30880, 18.31465…
$ lon     <dbl> -64.77668, -64.68200, -64.68392, -64.72294, -64.…
$ hab     <chr> "PVMT", "AGRF", "AGRF", "AGRF", "PVMT", "SCR", "…


2. How many samples are in the data set? 

```{r}
nrow(d)
```
2505

3. How many unique habitat types are there?  What are the names of each habitat?

```{r}
unique(d$hab)
```
[1] "PVMT" "AGRF" "SCR"  "PTRF" "BDRK" "HARD"
6

4. How many sites are there in the "PVMT" habitat type?
```{r}
d |> filter(hab == "PVMT")
```
571


5. How many sites are there in the "PVMT" or "AGRF" habitat types?
```{r}
d |> filter(hab=="PVMT" | hab=="AGRF")
```
1538

6. How many sites are there in all the habitat types EXCEPT "PVMT"?
d |> filter(hab!="PVMT")
1934

7. Show the first 5 records (rows) of...
    - all sites sampled in "AGRF" or "SCR"
    - only show the "yr_site" and "hab" columns

d |> 
filter(hab=="AGRF" | hab=="SCR") |> 
head()
# A tibble: 6 × 5
  yr_site   year   lat   lon hab  
  <chr>    <dbl> <dbl> <dbl> <chr>
1 2001_243  2001  18.3 -64.7 AGRF 
2 2001_244  2001  18.3 -64.7 AGRF 
3 2001_247  2001  18.3 -64.7 AGRF 
4 2001_249  2001  18.3 -64.7 SCR  
5 2001_256  2001  18.3 -64.7 AGRF 
6 2001_260  2001  18.3 -64.7 AGRF 

d |> 
filter(hab=="AGRF" | hab=="SCR") |> 
select(yr_site,hab) |> 
head()
# A tibble: 6 × 2
  yr_site  hab  
  <chr>    <chr>
1 2001_243 AGRF 
2 2001_244 AGRF 
3 2001_247 AGRF 
4 2001_249 SCR  
5 2001_256 AGRF 
6 2001_260 AGRF 

    
<!-- extra line break -->
<br>

8. What is the northern most site sampled in the "BDRK" habitat?

d |> 
filter(hab=="BDRK") |> 
slice_max(order_by=lat) |> 
pull(yr_site)
[1] "2019_5450"

9. What is the western most site sampled in either "AGRF" or "PTRF" habitat?

d |> 
filter(hab=="AGRF"|hab=="PTRF") |> 
slice_min(order_by=lon) |> 
pull(yr_site)
[1] "2015_9758"

10. How many years were surveyed in the data set?

length(unique(d$year))
16

11. What is the eastern most site sample in the "AGRF" habitat in 2004?

d |> 
filter(hab=="AGRF" & year==2004) |> 
slice_max(lon) |> 
pull(yr_site)
[1] "2004_1821"

12. Of the total number of samples, what percentage was surveyed on patch reef ("PTRF")?
mean(d$hab=="PTRF")*100
[1] 12.77445



